Enriching Object Detection with 2D-3D Registration and Continuous Viewpoint

### 2014-11-14 egpaper_final.tex

We propose Non-Zero Whitened Histograms of Orientations (NZ-WHO), that are both
more discriminative than WHO, and can be generated in 70ms (compared to several
seconds for WHO). This speed up means we can generate templates on the fly,
allowing our approach to evaluate arbitrary viewpoint hypotheses.

\scream{chris:We propose Non-Zero Whitened Histograms of Orientations (NZ-WHO),
an improvement over WHO which is both more discriminative than WHO and can be
computed many orders of magnitude faster than WHO (70ms as opposed to to several
seconds for WHO). This speedup allows us to generate NZ-WHO templates on the
fly, allowing us to quickly evaluate arbitrary viewpoint hypotheses.}

The computational efficiency and numerical stability of our approach comes from
using the iterative conjugate gradient method to whiten the HOG template.
Whitening a feature vector $x$ requires a system of linear equations $\Sigma w =
(x - \mu)$ to be solved, usually by decomposing the covariance matrix $\Sigma$
(with LU or Cholesky decomposition). Since the covariance matrices of HOG
templates have nice spectral properties (see Section \ref{sec:fast_whitening}),
the conjugate gradient method will converge to the solution several orders of
magnitude faster than decomposition methods.

@@

Table \ref{tab:who_initializations} provides a detailed comparison of WHO and
NZ-WHO templates, illustrating that our method is faster and more discriminative
than Whitened Histograms of Orientations. \scream{Does the table show this?}

% We compare the performance of WHO and NZ-WHO on 3D Object dataset \cite{Savarese07} and report performance on Table \ref{tab:who_initializations}.

@@

>>>>

\subsection{Overview}
Our method attempts to match CAD models

Our method is in line with \cite{Aubry14, Lim14} in the sense that we
use only renderings, not real images. However, rather than using fixed
sized mid-level patches, we use whole objects and make root
templates. By using root templates, there is no need to learn a
discriminative patch dictionary or to calibrate the patches and an
additional inference stage is not required. However, whitening the
whole template is prohibitive for several reasons. All templates have
different aspect ratios making it difficult to generate the covariance
matrix $\Sigma$. Since the root templates are larger than mid-level
patches, a very large auto-covariance $\Gamma$ is required. Lastly,
time complexity of decorrelating a template with $n$ cells increases
as $O(n^3)$, resulting in slow template generation and large residuals
due to numerical errors accumulated while decomposing the covariance
matrix $\Sigma$.

To overcome these difficulties, we propose the Non-Zero Whitened
Histograms of Orientations (NZ-WHO) with iterative Conjugate Gradient to
generate templates in real time and enrich rectangular boxes with high
quality 2D-3D matching. The 2D-3D matching provides a CAD model,
sub-category, rendering, depth, viewpoint and focal length.

Our method can work as a standalone 2D-3D matching algorithm, or even
as a standalone ensemble of exemplar-based object detectors (Figure
\ref{fig:front}) but its real power comes from combining generic
object detectors and complement rectangular bounding boxes. Our method
is simple in design; is several orders of magnitude faster than
traditional method; requires only CAD models; does not need neither images nor
calibration, yet is powerful in giving high-quality metadata. These
properties make our method ideal for complementing object detections.

To complement detection bounding boxes, we first search possible 2D-3D
matching by running cached discrete viewpoint templates. Then, for the
most probable match, we refine the initial matching using Metropolis
Hastings. Since it requires iterations of proposal and validation, we
present real time analysis of NZ-WHO template generation to show that
this can be done on-the-fly. The fine tuning stage process jointly
optimizes translation, yaw, roll, pitch, scale, and focal length
continuously and also searches for different CAD models.

We propose Non-Zero Whitened Histograms of Orientations (NZ-WHO), that are both more discriminative than WHO, and can be generated in 70ms (compared to several seconds for WHO). This speed up means we can generate templates on the fly, allowing our approach to evaluate arbitrary viewpoint hypotheses. 
\scream{chris:We propose Non-Zero Whitened Histograms of Orientations (NZ-WHO), an improvement over WHO which is both more discriminative than WHO and can be computed many orders of magnitude faster than WHO (70ms as opposed to to several seconds for WHO). This speedup allows us to generate NZ-WHO templates on the fly, allowing us to quickly evaluate arbitrary viewpoint hypotheses.}

The computational efficiency of our approach comes from using the iterative conjugate gradient method to whiten the HOG template. Whitening a feature vector $x$ requires a system of linear equations $\Sigma w = (x - \mu)$ to be solved, usually by decomposing the covariance matrix $\Sigma$ (with LU or Cholesky decomposition). Since the covariance matrices of HOG templates have nice spectral properties (see Section \ref{sec:fast_whitening}), the conjugate gradient method will converge to the solution several orders of magnitude faster than decomposition methods.

@@

Table \ref{tab:who_initializations} provides a detailed comparison of WHO and NZ-WHO templates, illustrating that our method is faster and more discriminative than Whitened Histograms of Orientations. \scream{Does the table show this?}
% We compare the performance of WHO and NZ-WHO on 3D Object dataset \cite{Savarese07} and report performance on Table \ref{tab:who_initializations}.

@@

Our method begins by convolving the
